---
title: 'Project Proposal'
author: 'Isaac Cheong, Sam Tan, Max Zhang'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

load('ICPSR_20520/DS0001/20520-0001-Data.rda')  # creates df called da20520.0001
cils <- da20520.0001
```

# Potential title: 
- Social mobility of the children of immigrants in America
- The Adaptation of the Immigrant Second Generation in America

# Introduction

What makes some second generation Americans more adaptive to the mainstrain society than others? In this report, we set out to investiage this question by predicting one of the important measurements for overall happiness: income and financial satisfication. 

Using data provided by the Center for Migration and Development at princeton univeristy, we will predict the income level of 2500+ second generation Americans residing in San Diego and Miami in 2005 based on factors such as their home environment, parents' language compentency, and their personal beliefs. 

# Data Description

# EDA

# Model Selection

EDA 
Shrikage? A way to reduce the variance of predictions by reducing the size of each coeffient, but the coeffients will be biased. LASSO 
Stepwise/forward/Backward selection? 
exhaustive search


# Model Diagnostics


# Prediction

# Conclusion

\newpage

# Appendix

```{r}
head(cils[ ,1:5], 5)
```




